SOMETHING ABOUT THE DISTRIBUTED AVERAGING BASED APPROXIMATION IN SENSOR NETWORKS

### 2022-10-10 paper.tex
transmission cost
intro

Main contribution: Avoid broadcasting or relaying of norm values, with distributed averaging approach and low iteration count.
Algorithms using a similar approach: \cite{yuDistributedBlindSystem2014, liuDistributedBlindIdentification2016,liuDistributedRecursiveBlind2017}, however these use high numbers of iterations (30) which leads to many per-frame transmissions.
Argument against broadcasting: (1)Large-scale networks, where not all nodes can physically communicate with each other, but that is not really that relevant for room acoustics... (2) just because. this is the network topology and configuratinos that was chosen as in a lot of literature.

Important context in introduction: BSI, WSN, topology, connection matrix,

What to show:
\begin{itemize}
  \item Algorithm extension
  \item Results
  \begin{itemize}
    \item compare results of unextended algorithm (optimal case)
    \item with extended: varied iteration count (more iterations: better result, BUT also even 1 iteration can lead to good results)
    \item varied mixing factor (there is some optimal \(\gamma\) apparently)
    \item Show transmission cost: transmission cost over number of nodes
  \end{itemize}
\end{itemize}

\section{DISTRIBUTED ADAPTIVE BLIND SYSTEM IDENTIFICATION}
\label{sec:dbsi}
Introduction of the algorithm from iwaenc paper. \cite{blochbergerDBSI}
\begin{align}
  \wf_i^{k+1} &= \underset{\wf_i}{\operatorname{argmin}} \, \mathcal{L}_{\rho} (\wf,\hf^k,\uuf^k)\label{eq:general_consensus_admm:local}\\
  \hf^{k+1} &= \underset{\hf, \|\hf\| = 1}{\operatorname{argmin}}\, \mathcal{L}_{\rho} (\wf^{k+1},\hf,\uuf^k)\label{eq:general_consensus_admm:global}\\
  \uuf_i^{k+1} &= \uuf_i^{k} + \rho \left( \wf_i^{k+1} - \tilde{\hf}_i^{k+1} \right),\label{eq:general_consensus_admm:dual}
\end{align}

The minimization problem for the local variable \(\wf_i\) \eqref{eq:general_consensus_admm:local} can be solved by various algorithms, in this case however we perform the adaptive update step as
\begin{equation}
    \wf_i^{m+1} = \wf_i^{m} - \mu \bm{{V}}_i^m \left( \hat{\aRhof}_i^m \wf_i^m + \uuf_i^m + \rho\left(\wf_i^m - \tilde{\hf}_i^{m}\right)\right),\label{eq:online_admm:local_update}
\end{equation}
where \(\mu\), (\(0  < \mu\leq 1\)), is a step size and \(\bm{{V}}_i^m = \left(\hat{\aRhof}_i^m + \rho \I \right)^{-1}\) is the regularized inverse Hessian of the problem.
The implicit overlap-save matrices in \(\bm{{V}}_i^m\) make this a constrained update step which is costly to compute, so in order to reduce computational complexity, we introduce the approximation
\begin{equation}
    \hat{\bm{{V}}}_i^m = \operatorname{diag} \left\{ \operatorname{diag} \left\{ \hat{\aRhof}_i^m \right\} + \rho \mtxb{1} \right\}^{-1},
\end{equation}
which is straightforward to compute.
It may be readily verified (cf. \cite{boydDistributedOptimizationStatistical2011}) that the solution to \eqref{eq:general_consensus_admm:global} is given by an update step which computes
\begin{equation}
    \hf_i^{m+1} = \frac{\bar{\hf}_i^{m+1}}{\sqrt{\sum_{j \in \Mset} \|\bar{\hf}_j^{m+1}\|^2}}\label{eq:online_admm:consensus_update}
\end{equation}
for each node \(i \in \Mset\) with the local unnormalized consensus
\begin{equation}
    \bar{\hf}_i^{m+1} = \bar{\wf}_i^{m+1} + \frac{1}{\rho} \bar{\uuf}_i^{m}.
\end{equation}
Here, \(\bar{\wf}_i = \frac{1}{N_i} \sum_{j \in \Tset_i} (\wf_j)_i\) and \(\bar{\uuf}_i = \frac{1}{N_i} \sum_{j \in \Tset_i} (\uuf_j)_i\), with \(N_i = | \Tset_i |\), are neigborhood averages of the channel response estimate and the dual variable of channel \(i\) respectively, computed at node \(i\). The values of all \(\|\bar{\hf}_i^{m+1}\|^2\), \(i \in \Mset\) are transmitted and relayed through the network until all nodes can compute the denominator in \eqref{eq:online_admm:consensus_update}.
Finally, the update of the dual variables is given by
\begin{equation}
    \uuf_i^{m+1} = \uuf_i^{m} + \rho \left( \wf_i^{m+1} - \tilde{\hf}_i^{m+1} \right).\label{eq:online_admm:dual_update}
\end{equation}

@@

\section{Adaptive Norm Estimation}
\label{sec:adaptivenormest}
Distributed averaging:
\begin{mini}{}{\|\W - \bm{1}\bm{1}^\T/n\|_2} {\label{eq1}}{}
  \addConstraint{\W}{\in \mathcal{C}}
  \addConstraint{\bm{1}^\T \W}{= \bm{1}^\T}
  \addConstraint{\W \bm{1}}{= \bm{1}}
\end{mini}
gives us the optimal weights \(\mtxb{W}_{opt}\). The distributed averaging approach is an iterative approach, however here, we will try to find a way to limit the number of iterations necessary. This goal comes from the fact that the iterations require data from neighboring nodes leading to additional transmissions (per frame).

We take avantage of the fact that the squared norm of the stacked solution vector can be separated into
\begin{equation}
  \| \h \|^2 = \sum_{j \in \Mset} \|\h_j\|^2.
\end{equation}
Then, we define the approximation of the global norm \(\hat{\phi}_i \approx \|\h\|^2\) at each node \(i\) at frame \(m\) as follows:
\begin{align}
  \phi_i^1 &= \hat{\phi}_i^{m}\\
  \phi_i^{k+1} &= \sum_{j \in \Tset_i} W_{opt,ij} \phi_j^{k}\label{eq:dist_norm_est_iter}\\
  \hat{\phi}_i^{m+1} &= \gamma \| \bar{\h}_j^{m+1} \|^2 + (1-\gamma) \phi_i^K
\end{align}
where \(\gamma\) is a exponential smoothing factor balancing new norm data with previous iterative estimate and \eqref{eq:dist_norm_est_iter} is applied \(K\) times.

The special case of \(K=1\) leads to the update step:
\begin{equation}
  \hat{\phi}_i^{m+1} = \gamma \| \bar{\h}_j^{m+1} \|^2 + (1-\gamma) \sum_{j \in \Tset_i} W_{opt,ij} \hat{\phi}_j^m
\end{equation}

We use these in an inbetween step to estimate the norms of the channel impulse responses recursively. The full algorithm then looks like
\begin{align}
  \w_i^{m+1} &= \operatorname{argmin} \mathcal{L}_\rho(\w,\h,\uu)\\
  \bar{\h}_i^{m+1} &= \bar{\w}_i^{m+1} + \frac{1}{\rho} \bar{\uu}_i^m\\
  \phi_i^1 &= \hat{\phi}_i^{m}\\
  \phi_i^{k+1} &= \sum_{j \in \Tset_i} W_{opt,ij} \phi_j^{k} \qquad\text{for } k=1...K\\
  \hat{\phi}_i^{m+1} &= \gamma \| \bar{\h}_j^{m+1} \|^2 + (1-\gamma) \phi_i^K\\
  \h_i^{m+1} &= \frac{\bar{\h}_i^{m+1}}{\sqrt{\hat{\phi}_i^K\,M}}\\
  \uu_i^{m+1} &= \uu_i^{m} + \rho \left(\w_i^{m+1} - \tilde{\h}_i^{m+1}\right).
\end{align}

\section{Transmission cost}
\label{sec:transcost}
Show effect on number of transmissions. Scale up to high numbers of nudes so neigborhoods \(<<\) entire network.

>>>>

Distributed algorithms have been an active area of research for quite some time, with numerous control, optimization, and signal processing applications.
With the ever-growing number of smart multimedia devices in today's surroundings providing ubiquitous processing and communication capabilities, distributed audio and speech signal processing also found their way into the spotlight.
Algorithms for distributed signal estimation \cite{5483092}, noise control and echo cancellation \cite{9670697}, as well as beamforming \cite{6663655,6329934,MARKOVICHGOLAN20154} amongst others, have been proposed.
The task of distributed single-input-multiple-output (SIMO) blind system identification (BSI) had contributions such as the adaptive cross-relation-based (CR) \cite{yuDistributedBlindSystem2014, liuDistributedBlindIdentification2016}.
Furthermore, we recently introduced an adaptive CR-based algorithm \cite{blochbergerDBSI} using the alternating direction method of multipliers (ADMM) \cite{boydDistributedOptimizationStatistical2011}.
All of the mentioned distributed BSI algorithms rely on information shared between neighboring sensor nodes within the network.
However, the CR-based BSI task necessitates a non-triviality constraint on the full system to be identified (we refer the reader to e.g. \cite{huangAdaptiveMultichannelLeast2002,huangClassFrequencydomainAdaptive2003}), which manifests itself as one or more global variables.
In this case, a global variable is a variable, the computation of which requires information from all nodes within the network.
To overcome the need for the network to be fully connected, \cite{yuDistributedBlindSystem2014, liuDistributedBlindIdentification2016} use an average consensus \cite{xiaoFastLinearIterations2004} approach where a secondary recursion estimates the global variable for each signal frame.
The algorithm in \cite{blochbergerDBSI} relies on node-wise values being relayed throughout the network.
Both approaches introduce additional transmissions of variables between nodes, the number of which, depending on network and neighborhood size, can be substantial and even unfeasable.

In this paper, we extend \cite{blochbergerDBSI} with a distributed averaging-based \cite{xiaoFastLinearIterations2004} estimation scheme for the global variable with the introduction of a mixing factor to include instantaneous values into the averaging recursion.
The mixing factor allows us (i) to reduce the number of secondary iterations significantly (down to 1) and (ii) track time-varying systems.
Simulation results show how the extension leads to BSI performance close to an optimal case where all information is available without the need of broadcasting variables to all nodes or a large number of estimation iterations.

% Main contribution: Avoid broadcasting or relaying of norm values, with distributed averaging approach and low iteration count.
% Algorithms using a similar approach: \cite{yuDistributedBlindSystem2014, liuDistributedBlindIdentification2016,liuDistributedRecursiveBlind2017}, however these use high numbers of iterations (30) which leads to many per-frame transmissions.
% Argument against broadcasting: (1)Large-scale networks, where not all nodes can physically communicate with each other, but that is not really that relevant for room acoustics... (2) just because. this is the network topology and configuratinos that was chosen as in a lot of literature.

% Important context in introduction: BSI, WSN, topology, connection matrix,

% What to show:
% \begin{itemize}
%   \item Algorithm extension
%   \item Results
%   \begin{itemize}
%     \item compare results of unextended algorithm (optimal case)
%     \item with extended: varied iteration count (more iterations: better result, BUT also even 1 iteration can lead to good results)
%     \item varied mixing factor (there is some optimal \(\gamma\) apparently)
%     \item Show transmission cost: transmission cost over number of nodes
%   \end{itemize}
% \end{itemize}

\section{DISTRIBUTED ADAPTIVE BLIND SYSTEM IDENTIFICATION USING ONLINE-ADMM}
\label{sec:dbsi}
In \cite{blochbergerDBSI}, we proposed an adaptive SIMO BSI algorithm based on Online-ADMM.
In this section, we will introduce parts which are necessary for the reader to understand the extension.
For the full algorithm, we refer the reader to the previous publication.

The coss-realtion problem 
\begin{align}
  \wf_i^{k+1} &= \underset{\wf_i}{\operatorname{argmin}} \, \mathcal{L}_{\rho} (\wf,\hf^k,\uuf^k)\label{eq:general_consensus_admm:local}\\
  \hf^{k+1} &= \underset{\hf, \|\hf\| = 1}{\operatorname{argmin}}\, \mathcal{L}_{\rho} (\wf^{k+1},\hf,\uuf^k)\label{eq:general_consensus_admm:global}\\
  \uuf_i^{k+1} &= \uuf_i^{k} + \rho \left( \wf_i^{k+1} - \tilde{\hf}_i^{k+1} \right),\label{eq:general_consensus_admm:dual}
\end{align}

% The minimization problem for the local variable \(\wf_i\) \eqref{eq:general_consensus_admm:local} can be solved by various algorithms, in this case however we perform the adaptive update step as
% \begin{equation}
%     \wf_i^{m+1} = \wf_i^{m} - \mu \bm{{V}}_i^m \left( \hat{\aRhof}_i^m \wf_i^m + \uuf_i^m + \rho\left(\wf_i^m - \tilde{\hf}_i^{m}\right)\right),\label{eq:online_admm:local_update}
% \end{equation}
% where \(\mu\), (\(0  < \mu\leq 1\)), is a step size and \(\bm{{V}}_i^m = \left(\hat{\aRhof}_i^m + \rho \I \right)^{-1}\) is the regularized inverse Hessian of the problem.
% The implicit overlap-save matrices in \(\bm{{V}}_i^m\) make this a constrained update step which is costly to compute, so in order to reduce computational complexity, we introduce the approximation
% \begin{equation}
%     \hat{\bm{{V}}}_i^m = \operatorname{diag} \left\{ \operatorname{diag} \left\{ \hat{\aRhof}_i^m \right\} + \rho \mtxb{1} \right\}^{-1},
% \end{equation}
% which is straightforward to compute.
It may be readily verified (cf. \cite{boydDistributedOptimizationStatistical2011}) that the solution to \eqref{eq:general_consensus_admm:global} is given by an update step which computes
\begin{equation}
    \hf_i^{m+1} = \frac{\bar{\hf}_i^{m+1}}{\sqrt{\sum_{j \in \Mset} \|\bar{\hf}_j^{m+1}\|^2}}\label{eq:online_admm:consensus_update}
\end{equation}
for each node \(i \in \Mset\) with the local unnormalized consensus
\begin{equation}
    \bar{\hf}_i^{m+1} = \bar{\wf}_i^{m+1} + \frac{1}{\rho} \bar{\uuf}_i^{m}.
\end{equation}
Here, \(\bar{\wf}_i = \frac{1}{N_i} \sum_{j \in \Tset_i} (\wf_j)_i\) and \(\bar{\uuf}_i = \frac{1}{N_i} \sum_{j \in \Tset_i} (\uuf_j)_i\), with \(N_i = | \Tset_i |\), are neigborhood averages of the channel response estimate and the dual variable of channel \(i\) respectively, computed at node \(i\). The values of all \(\|\bar{\hf}_i^{m+1}\|^2\), \(i \in \Mset\) are transmitted and relayed through the network until all nodes can compute the denominator in \eqref{eq:online_admm:consensus_update}.
% Finally, the update of the dual variables is given by
% \begin{equation}
%     \uuf_i^{m+1} = \uuf_i^{m} + \rho \left( \wf_i^{m+1} - \tilde{\hf}_i^{m+1} \right).\label{eq:online_admm:dual_update}
% \end{equation}

@@

\section{Adaptive Norm Estimation}
\label{sec:adaptivenormest}
Distributed averaging:
\begin{mini}{}{\|\W - \bm{1}\bm{1}^\T/n\|_2} {\label{eq1}}{}
  \addConstraint{\W}{\in \mathcal{C}}
  \addConstraint{\bm{1}^\T \W}{= \bm{1}^\T}
  \addConstraint{\W \bm{1}}{= \bm{1}}
\end{mini}
gives us the optimal weights \(\mtxb{W}_{o}\) where the weights used by node \(i\) to combine values from neighbor nodes \(j \in \Rset_i\) are given at the \(i\)-th row. The distributed averaging approach is an iterative approach, however here, we will try to find a way to limit the number of iterations necessary. This goal comes from the fact that the iterations require data from neighboring nodes leading to additional transmissions (per frame).

% We take avantage of the fact that the squared norm of the stacked solution vector can be separated into
% \begin{equation}
%   \| \h \|^2 = \sum_{j \in \Mset} \|\h_j\|^2.
% \end{equation}
We define the approximation of the global norm \(\hat{\phi}_i \approx \|\h\|^2\) at each node \(i\) at frame \(m\) as follows:
\begin{equation}
  \phi_i^{m,k+1} = \sum_{j \in \Rset_i} w_{ij} \phi_j^{m,k}\quad\text{for }k=1,\ldots,K\label{eq:dist_norm_est_iter}
\end{equation}
with the initial value at \(m=1,k=1\) defined as \(\phi_i^{1,1}=\|\h_i^{1}\|^2\).
\eqref{eq:dist_norm_est_iter} is applied \(K\) times.
We now define the frame estimate of the global squared norm as
\begin{equation}
  \hat{\phi}_i^{m+1} = \gamma_i \bar{\phi}_i^{m} + (1-\gamma) \phi_i^{m,K}
\end{equation}
where \(\bar{\phi}_i^{m} = \sum_{j \in \Rset_i} w_{ij} \|\h_i^m\|^2\) is an instantaneous weighted neighborhood average of node-wise squared norm values and \(\gamma_i\) is a mixing factor.
This factor can be chosen as a fixed value, where simulations have shown that values around \(\gamma_i = 0.02\) lead to satisfying results (cf. FIGURE) or adaptively.
For this, we use the instantaneous values to track stationarity of the squared norms by setting the factor \(\gamma_i^{m}\) proportional to \(| \bar{\phi}_i^{m} - \bar{\phi}_i^{m-1} |\), the absolute difference between instantaneous neighborhood norm values of subsequent frames.

The special case of \(K=1\) leads to the update step:
\begin{equation}
  \hat{\phi}_i^{m+1} = \gamma_i \sum_{j \in \Rset_i} w_{ij} \| \bar{\h}_j^{m+1} \|^2 + (1-\gamma_i) \sum_{j \in \Rset_i} w_{ij} \hat{\phi}_j^m
\end{equation}

We use these in an inbetween step to estimate the norms of the channel impulse responses recursively. The full algorithm then looks like

\begin{algorithm}
  \caption{Estimation of norm values}\label{alg:davg_norm_est}
  $i\gets 10$\;
  \eIf{$i\geq 5$}
  {
      $i\gets i-1$\;
  }{
      \If{$i\leq 3$}
      {
          $i\gets i+2$\;
      }
  }
\end{algorithm}

% \begin{align}
%   \w_i^{m+1} &= \operatorname{argmin} \mathcal{L}_\rho(\w,\h,\uu)\\
%   \bar{\h}_i^{m+1} &= \bar{\w}_i^{m+1} + \frac{1}{\rho} \bar{\uu}_i^m\\
%   \phi_i^1 &= \hat{\phi}_i^{m}\\
%   \phi_i^{k+1} &= \sum_{j \in \Rset_i} w_{ij} \phi_j^{k} \qquad\text{for } k=1,\ldots,K\\
%   \hat{\phi}_i^{m+1} &= \gamma \| \bar{\h}_j^{m+1} \|^2 + (1-\gamma) \phi_i^K\\
%   \h_i^{m+1} &= \frac{\bar{\h}_i^{m+1}}{\sqrt{\hat{\phi}_i^K\,M}}\\
%   \uu_i^{m+1} &= \uu_i^{m} + \rho \left(\w_i^{m+1} - \tilde{\h}_i^{m+1}\right).
% \end{align}

\section{Transmission cost}
\label{sec:transcost}
Show effect on number of transmissions. Scale up to high numbers of nodes so neigborhoods \(<<\) entire network.

\begin{align}
  \mathcal{O}(&(M-1)^2) & &+ & \mathcal{O}(&(M-1)^2)\\
  \mathcal{O}(&M-1) & &+ & \mathcal{O}(&(M-1)^2)\\
  \mathcal{O}(&M \bar{N}_i R) & &+ & \mathcal{O}(&M \bar{N}_i R)
\end{align}

Or per node:
\begin{align}
  \mathcal{O}(&M-1) & &+ & \mathcal{O}(&M-1)\\
  \mathcal{O}(&1) & &+ & \mathcal{O}(&M-1)\\
  \mathcal{O}(&\bar{N}_i R) & &+ & \mathcal{O}(&\bar{N}_i R)
\end{align}

\begin{figure}
  \centering
  \input{simulations/plots/transcost/transcost.pgf}
  \input{simulations/plots/transcostnode/transcostnode.pgf}
  \vspace*{-0.8cm}
  \caption[]{The order of transmission cost of entire network or single node (choose which to keep if at all.).}
  \label{fig:transcost:bigo}
\end{figure}
